### 2022-07-06 manuscript.Rmd
Update for re-submission

author:
  - name          : "Alexander Enge"
    affiliation   : "1,2,*"
  - name          : "Shreya Kapoor"
    affiliation   : "1"
  - name          : "Anne-Sophie Kieslinger"
    affiliation   : "1"
  - name          : "Michael A. Skeide"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences, Stephanstraße 1a, 04103 Leipzig, Germany"
  - id            : "2"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin, Rudower Chaussee 18, 12489 Berlin, Germany"
  - id            : "*"
    institution   : "Corresponding author (enge@cbs.mpg.de)"

bibliography      : "manuscript_files/references.bib"
csl               : "manuscript_files/pnas.csl"

@@

header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage{lineno}
  - \usepackage{makecell}
  - \usepackage{setspace}
  - \DeclareCaptionLabelSeparator{bar}{ | }
  - \DeclareCaptionTextFormat{tabletext}{\textup{#1}}
  - \captionsetup[table]{belowskip=20pt,font=small,labelfont=bf,labelsep=bar,textfont=bf,textformat=tabletext}
  - \captionsetup[figure]{belowskip=10pt,font=small,labelfont=bf,labelsep=bar,name=Fig.,textfont=normalfont}
---

@@

\linenumbers

# Abstract

Mental rotation, the cognitive process of moving an object in mind to predict how it looks in a new orientation, is tightly coupled to intelligence, learning, and educational achievement [@shepard1971; @hegarty1999; @johnson2005]. On average, males solve mental rotation tasks slightly faster than females [@linn1985; @voyer1995; @voyer2011; @maeda2013]. When such behavioral differences emerge during development, however, remains poorly understood [@lauer2019a]. Here we analyzed effect sizes derived from 59 experiments conducted in 1,798 infants aged 3--16 months. We robustly found that male infants recognized rotated objects more reliably than female infants. These findings indicate that gender differences in mental rotation are already present in the first months of life.

# Significance statement

Are there gender differences in cognitive ability? So far, putative differences in most cognitive domains turned out to be myths. However, mental rotation, the ability to move an object in mind to predict how it would look in a new orientation, is considered as a serious candidate. Here we investigated how early during development mental rotation behavior differs in male and female infants. We found that, on average, males recognized rotated objects more reliably than females as early as in the first months of life. These results are an important step towards narrowing down possible social and genetic origins of gender differences in mental rotation ability.

# Introduction

@@

Previous meta-analyses revealed that males solve mental rotation tasks slightly faster than females on average [@linn1985; @voyer1995; @voyer2011; @maeda2013]. Effect sizes of this difference, however, are heterogeneous and often only medium in size (mean weighted $g = 0.37\text{--}0.73$). Interestingly, a recent meta-analysis in 3--17-year-old children and adolescents suggests only a small-to-medium difference (mean weighted $g = 0.39$) [@lauer2019a]. Whether gender differences in mental rotation behavior already emerge during infancy remains unknown.

In the present study, we meta-analyzed `r descriptives$n_experiments` effect sizes derived from looking times in mental rotation tasks performed by `r formatC(descriptives$n_infants, big.mark = ",")` infants (`r print_perc(descriptives$percent_female)` female) aged `r descriptives$min_age_months` to `r descriptives$max_age_months` months (mean age of `r descriptives$mean_age_weighted`; Supplementary Table\ \@ref(tab:tabs1))[@kellman1984; @rochat1996; @hespos1997; @moore2008; @quinn2008; @moore2011; @frick2013b; @mohring2013; @schwarzer2013; @schwarzer2013a; @frick2014b; @quinn2014; @lauer2015; @antrilli2016; @christodoulou2016; @lauer2016; @constantinescu2018; @erdmann2018; @gerhard2018; @slone2018; @kaaz2020; @gerhard-samunda2021; @kelch2021]. All tasks were embedded either into habituation experiments (`r descriptives$n_experiments_habituation` experiments) or violation of expectation experiments (`r descriptives$n_experiments_voe` experiments) [@fantz1964; @baillargeon1985]. These experiments comprised real world stimuli (e.g., toy objects), three-dimensional digital stimuli (e.g., cube figures), or two-dimensional digital stimuli (e.g., digits). In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object or with the familiar object at a new angle. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

@@

For our first meta-analytic model, effect sizes were quantified as the standardized mean difference in infants' looking times for novel and familiar rotated objects. Using this effect size index, we ran a Bayesian three-level random-effects model to test if there was evidence that infants did perform mental rotation. Across studies, infants indeed looked longer at novel rotated objects than at familiar rotated objects, with a standardized mean difference of $g = `r summ_meta$intercept`$, 95% credible interval (CrI) [`r summ_meta$intercept.lower`, `r summ_meta$intercept.upper`] (Fig.\ \@ref(fig:fig1)). The probability for this effect being greater than zero was `r print_prob(hyp_meta$Post.Prob)`. The heterogeneity of effect sizes was $\sigma_{\text{experiment}}^2 = `r summ_meta$sigma2_experiment`$, 95% CrI [`r summ_meta$sigma2_experiment.lower`, `r summ_meta$sigma2_experiment.upper`], at the experiment level and $\sigma_{\text{article}}^2 = `r summ_meta$sigma2_article`$, 95% CrI [`r summ_meta$sigma2_article.lower`, `r summ_meta$sigma2_article.upper`] at the article level. Therefore, `r print_perc(1 - summ_meta$icc)` of the heterogeneity between effect sizes was attributable to differences between experiments *within* articles and `r print_perc(summ_meta$icc)` was attributable to differences *between* articles.

@@

As a next step, we conducted a meta-regression analysis to test if the gender of the infants, their age, or the type of mental rotation task was related to mental rotation performance (Fig.\ \@ref(fig:fig2)). Indeed, experiments with all-male samples revealed larger looking time differences than experiments with all-female samples, $b = `r summ_reg$male_female`$ (where $b = \Delta{g}$), 95% CrI [`r summ_reg$male_female.lower`, `r summ_reg$male_female.upper`]. The probability for this effect being larger than zero was `r print_prob(hyps_brm_reg$male_female$Post.Prob)`. We found no difference between mixed-gender and all-female samples, $b = `r -summ_reg$female_mixed`$, 95% CrI [`r -summ_reg$female_mixed.lower`, `r -summ_reg$female_mixed.upper`] (`r print_prob(hyps_brm_reg$female_mixed$Post.Prob)` probability of an effect larger than zero). Additionally, mean age was not related to mental rotation performance, with a change per year of $b = `r summ_reg$age`$, 95% [`r summ_reg$age.lower`, `r summ_reg$age.upper`]. We also did not detect an interaction between gender and age ([females - mixed] $\times$ age: $b = `r summ_reg$female_mixed_age`$, 95% CrI [`r summ_reg$female_mixed_age.lower`, `r summ_reg$female_mixed_age.upper`]; [males - females] $\times$ age: $b = `r summ_reg$male_female_age`$, 95% CrI [`r summ_reg$male_female_age.lower`, `r summ_reg$male_female_age.upper`]). Finally, there was weak evidence that violation of expectation tasks yielded larger effects than habituation tasks, $b = `r summ_reg$voe_habituation`$, 95% [`r summ_reg$voe_habituation.lower`, `r summ_reg$voe_habituation.upper`]. The probability for this effect being greater than zero was `r print_prob(hyps_brm_reg$voe_habituation$Post.Prob)`.

@@

# Combine the two regression plots
plot_grid(
  plot_reg, plot_coef,
  nrow = 1, rel_widths = c(3, 2), labels = "auto",
  label_size = 11, label_fontfamily = "Helvetica",
  label_x = 0.008, label_y = 0.99
) +
  draw_plot(plot_reg_legend, x = 0.385, y = 0.865, hjust = 0.5, vjust = 0.5)

@@

(ref:fig2-caption) **Effects of gender, age, and task.** **a,** Squares show the effect size (Hedges' $g$) for infants' mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 50% and 95% credible interval (CrI). **b,** Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 50% CrI (thick black lines) and 95% CrI (thin black lines). Gray curves indicate the posterior distribution for each effect.

@@

To confirm the gender difference between males and females, we set up another Bayesian meta-analysis, this time focusing on the looking-time contrasts between male and female infants reported *within* each experiment by the original authors. Our additional analysis revealed a meta-analytic effect size of $g = `r summ_gender$intercept`$, 95% CrI [`r summ_gender$intercept.lower`, `r summ_gender$intercept.upper`] (Supplementary Fig.\ \@ref(fig:figs1)) and a probability of this effect being greater than zero of `r print_prob(hyp_gender$Post.Prob)`. The heterogeneity of effect sizes was $\sigma_{\text{experiment}}^2 = `r summ_gender$sigma2_experiment`$, 95% CrI [`r summ_gender$sigma2_experiment.lower`, `r summ_gender$sigma2_experiment.upper`], at the experiment level and $\sigma_{\text{article}}^2 = `r summ_gender$sigma2_article`$, 95% CrI [`r summ_gender$sigma2_article.lower`, `r summ_gender$sigma2_article.upper`] at the article level. Therefore, `r print_perc(1 - summ_gender$icc)` of the heterogeneity between effect sizes was attributable to differences between experiments *within* articles and `r print_perc(summ_gender$icc)` was attributable to differences *between* articles.

@@

We inspected funnel plots and performed Egger regression tests to examine the possibility of publication bias in the literature included here. For the meta-analysis of mental rotation performance (Fig.\ \@ref(fig:fig3)a), we observed a slight asymmetry in the funnel plot, indicating a small publication bias. This was confirmed by an Egger regression test that showed a reliable association between effect sizes and their corresponding standard errors, $b = `r coef(egger_meta$fit)["Xsei"]`$, $t(`r egger_meta$dfs`) = `r egger_meta$zval`$, $P = `r print_num(egger_meta$pval, digits = 3)`$, 95% confidence interval (CI) [`r egger_ci_meta["2.5 %"]`, `r egger_ci_meta["97.5 %"]`] (two-sided test). Nevertheless, a jackknife (leave-one-out) analysis confirmed that the current results are robust to the effects of individual outlier experiments (Supplementary Table\ \@ref(tab:tabs3)). For the meta-analysis of gender differences in mental rotation performance (Fig.\ \@ref(fig:fig3)b), the asymmetry in the funnel plot was less pronounced and the slope of the Egger regression test was not statistically significant, $b = `r coef(egger_gender$fit)["Xsei"]`$, $t(`r egger_gender$dfs`) = `r egger_gender$zval`$, $P = `r print_num(egger_gender$pval, digits = 3)`$, 95% CI [`r egger_ci_gender["2.5 %"]`, `r egger_ci_gender["97.5 %"]`] (two-sided test).

@@

# Combine plots and legend
plot_grid(
  plotlist = plots_funnel, nrow = 1, labels = "auto",
  label_size = 11, label_fontfamily = "Helvetica"
) +
  draw_plot(legend_funnel, x = -0.39, y = 0.355)

@@

(ref:fig3-caption) **Evaluation of publication bias.**  Funnel plots for the meta-analysis of mental rotation performance (**a**) and for the meta-analysis of gender differences in mental rotation performance (**b**). The plots show the standard error and effect size for each of the individual experiments as a black square. The funnel contours (diagonal black lines) depict a 95% pseudo-confidence interval around the meta-analytic effect size (vertical black line). Gray shades indicate a 95% pseudo-confidence interval (dark gray) and a 99% pseudo-confidence interval (light gray) under the null hypothesis. These shades thus illustrate which of the original experiments observed a significant effect. For the meta-analysis of mental rotation performance, a slight asymmetry induced by the underrepresentation of experiments with high standard errors and small effect sizes suggests a small publication bias.

@@

To facilitate the interpretability of our models, the effect size of the looking time difference between novel and familiar rotation events was coded in a linear fashion. Accordingly, a positive looking time difference (i.e., a novelty preference) was taken as stronger evidence for mental rotation than a looking time difference of zero (i.e., no systematic preference). A zero difference in turn was taken as stronger evidence for mental rotation than a negative looking time difference (i.e., a familiarity preference). This approach can be corroborated by the established view that novelty preference is generally considered as the paradigmatic mental rotation behavior (for related discussion, see [@houston-price2004; @black2017; @cristia2018]). Nevertheless, some authors have argued that familiarity preference should also be taken as evidence for mental rotation ability [@erdmann2018; @moore2020; @johnson2020]. Interpreting novelty preference *and* familiarity preference as successful mental rotation would have increased the effect sizes reported here.

@@

A number of additional factors have been associated with individual differences in infant mental rotation performance. For example, mental rotation is related to previous relevant experience with the particular objects used in the specific task [@mohring2013; @schwarzer2013a; @slone2018]. This relation also applies to previous experience with manually rotating toys [@schwarzer2013a]. While these preliminary results require replication, they are in line with the longstanding notion that prior knowledge is the strongest predictor of learning outcomes in a range of cognitive domains (e.g., [@ausubel1968; @bradley1983; @halberda2008]). Furthermore, there is yet to be confirmed preliminary evidence for possible links between mental rotation performance and several sensory motor skills including fine and gross motor skills, oculomotor control, and crawling skills [@schwarzer2013; @schwarzer2013a].

@@

# Materials and methods

@@

In the present meta-analysis, we followed the established PRISMA 2020 guidelines (Preferred Reporting Items for Systematic reviews and Meta-Analyses) [@page2021]. The PRISMA checklist is provided in Supplementary Table\ \@ref(tab:tabs4) while Supplementary Fig.\ \@ref(fig:figs2) displays the PRISMA flowchart.

@@

We entered the search terms ("mental rotation" OR "mental transformation" OR "spatial rotation" OR "spatial transformation" OR "spatial ability" OR "spatial skills") AND ("infant" OR "infants" OR "infanthood" OR "toddler" OR "toddlers" OR "toddlerhood" OR "child" OR "children" OR "childhood" OR "month" OR "months") into four online databases (APA PsycINFO, PubMed/MEDLINE, Scopus, and ProQuest Dissertations & Theses Global). All database queries were completed on December 6, 2021. We configured the databases to check for article titles, abstracts, and keywords while applying no other filters or limits. This yielded 2,616 articles in total, 1,954 of which remained after removing duplicate records (Supplementary Fig.\ \@ref(fig:figs2)). We further identified 76 articles by screening the reference sections of previous reviews and meta-analyses on mental rotation and related skills [@frick2014; @johnson2020; @kubicek2018; @lauer2019a; @linn1985; @moore2020; @uttal2013; @voyer1995; @yang2020]. Of these, 34 articles had not been covered by the database search. We also identified 94 articles by screening the reference sections of all publications that had been included after the first pass of the selection process. Of these, 49 articles had not been covered by the database search. Accordingly, we screened 2,037 unique articles in total.

@@

Two independent raters read the abstract and, if necessary, relevant sections of the full text to check if an article fulfilled the inclusion criteria. Interrater agreement for the binary decision to include versus exclude an article was `r print_perc(irr_bin_perc)` ($\kappa_w$ [Cohen's weighted kappa] $= `r irr_bin_kappa_w$estimate`$, 95% CI [`r irr_bin_kappa_w$lower`, `r irr_bin_kappa_w$upper`]). Interrater agreement for the specific eligibility criteria was `r print_perc(irr_code_perc)` ($\kappa_w$ $= `r irr_code_kappa_w$estimate`$, 95% CI [`r irr_code_kappa_w$lower`, `r print_num(irr_code_kappa_w$upper)`]). Cases where the two ratings diverged were resolved via discussion among all raters until a consensus was reached. One article [@mash2007] was excluded because the authors used a unique mental rotation paradigm that was not comparable to the paradigms used in the other articles. Another article [@pedrett2020] was excluded because the average age of the infants studied (`r print_num(descs_pedrett2020$age_months, digits = 1)` months) was almost twice as high as the average age of the next article (`r print_num(descs_pedrett2020$next_largest_age_months, digits = 1)` months; $Z = `r descs_pedrett2020$z`$ compared to all articles). We therefore decided to narrow our analysis from the first 3 years of life to the first 16 months of life. This procedure led to a total of `r descriptives$n_articles` articles being included in the meta-analysis (Supplementary Fig.\ \@ref(fig:figs2)).

@@

We verified the results obtained from our Bayesian analyses using classical frequentist meta-analysis and meta-regression. To this end, we used the metafor package (Version `r packageVersion("metafor")`) [@viechtbauer2010] in R to specify the same three-level models as described above but without the Bayesian priors (Supplementary Table\ \@ref(tab:tabs2)). These models were fitted using restricted maximum likelihood estimation (REML). To verify that this procedure converged on the most probable estimates, we examined profile likelihood plots [@raue2009; @viechtbauer2010] for the two variance components in the model (i.e., the random effects of experiments and articles; Supplementary Fig.\ \@ref(fig:figs4)).

>>>>

author:
  - name          : "Alexander Enge"
    affiliation   : "1,2"
    corresponding : "yes"
    address       : "Stephanstraße 1a"
    email         : "enge@cbs.mpg.de"
  - name          : "Shreya Kapoor"
    affiliation   : "1"
  - name          : "Anne-Sophie Kieslinger"
    affiliation   : "1"
  - name          : "Michael A. Skeide"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Research Group Learning in Early Childhood, Max Planck Institute for Human Cognitive and Brain Sciences, Leipzig, Germany"
  - id            : "2"
    institution   : "Department of Psychology, Humboldt-Universität zu Berlin"

authornote: |
  We have no conflicts of interest to disclose.
  The data and code for this study are available at [https://osf.io/k3wdg/?view_only=e3b345f00ee84d318a9990689a1c5a9e](https://osf.io/k3wdg/?view_only=e3b345f00ee84d318a9990689a1c5a9e).

abstract: |
  Mental rotation, the cognitive process of moving an object in mind to predict how it looks in a new orientation, is tightly coupled to intelligence, learning, and educational achievement [@shepard1971; @hegarty1999; @johnson2005]. On average, males solve mental rotation tasks slightly faster than females [@linn1985; @voyer1995; @voyer2011; @maeda2013]. When such behavioral differences emerge during development, however, remains poorly understood [@lauer2019a]. Here we analyzed effect sizes derived from 59 experiments conducted in 1,798 infants aged 3--16 months. We robustly found that male infants recognized rotated objects more reliably than female infants. These findings indicate that gender differences in mental rotation are already present in the first months of life.
keywords: "mental rotation; spatial cognition; infants; gender; meta-analysis"

bibliography      : "manuscript_files/references.bib"
csl               : "manuscript_files/apa.csl"

@@

header-includes:
  - \geometry{a4paper}
  - \raggedbottom
  - \usepackage[all]{nowidow}
  - \usepackage{lineno}
  - \usepackage{makecell}
  - \usepackage{setspace}
  - \DeclareCaptionTextFormat{tabletext}{\textup{#1}}
  - \captionsetup[table]{belowskip=20pt,labelfont=bf,labelsep=period,textfont=bf,textformat=tabletext}
  - \captionsetup[figure]{belowskip=10pt,font=small,labelfont=bf,labelsep=period,textfont=normalfont}
---

@@

# Gender differences in infant mental rotation

@@

Previous meta-analyses revealed that males solve mental rotation tasks slightly faster than females on average [@linn1985; @voyer1995; @voyer2011; @maeda2013]. Effect sizes of this difference, however, are heterogeneous and often only medium in size (mean weighted $g = 0.37\text{--}0.73$). Interestingly, a recent meta-analysis in 3--17-year-old children and adolescents suggests only a small-to-medium difference [mean weighted $g = 0.39$\; @lauer2019a]. Whether gender differences in mental rotation behavior already emerge during infancy remains unknown.

In the present study, we meta-analyzed `r descriptives$n_experiments` effect sizes derived from looking times in mental rotation tasks performed by `r formatC(descriptives$n_infants, big.mark = ",")` infants (`r print_perc(descriptives$percent_female)` female) aged `r descriptives$min_age_months` to `r descriptives$max_age_months` months (mean age of `r descriptives$mean_age_weighted`\; Supplementary Table\ \@ref(tab:tabs1)) [@kellman1984; @rochat1996; @hespos1997; @moore2008; @quinn2008; @moore2011; @frick2013b; @mohring2013; @schwarzer2013; @schwarzer2013a; @frick2014b; @quinn2014; @lauer2015; @antrilli2016; @christodoulou2016; @lauer2016; @constantinescu2018; @erdmann2018; @gerhard2018; @slone2018; @kaaz2020; @gerhard-samunda2021; @kelch2021]. All tasks were embedded either into habituation experiments (`r descriptives$n_experiments_habituation` experiments) or violation of expectation experiments (`r descriptives$n_experiments_voe` experiments) [@fantz1964; @baillargeon1985]. These experiments comprised real world stimuli [e.g., toy objects\; @antrilli2016], three-dimensional digital stimuli [e.g., cube figures\; @moore2008], or two-dimensional digital stimuli [e.g., digits\; @quinn2008]. In habituation experiments, infants repeatedly saw an object until their looking times declined before they were presented with a mirror image of the object or with the familiar object at a new angle. Longer looking times at the mirror image were taken as evidence that an infant still recognized the familiar object after rotation through the new orientation. In violation of expectation experiments, infants were habituated to an object that was revolving repeatedly through a certain angle. Then, either the familiar object or an unseen object was shown while they were revolving through a certain angle. Subsequently, the object disappeared behind an occluder. Finally, the occluder was removed and either the object or its mirror object were shown. Larger differences in looking times for the familiar object versus the unseen mirror image (at the new angle) were taken as evidence that an infant still recognized the familiar object after rotation through the new angle. In contrast, looking times for both objects were expected to be similar if an infant did not recognize the familiar object from the new angle.

@@

For our first meta-analytic model, effect sizes were quantified as the standardized mean difference in infants' looking times for novel and familiar rotated objects. Using this effect size index, we ran a Bayesian three-level random-effects model to test if there was evidence that infants did perform mental rotation. Across studies, infants indeed looked longer at novel rotated objects than at familiar rotated objects, with a standardized mean difference of $g = `r summ_meta$intercept`$, 95% credible interval (CrI) [`r summ_meta$intercept.lower`, `r summ_meta$intercept.upper`] (Figure\ \@ref(fig:fig1)). The probability for this effect being greater than zero was `r print_prob(hyp_meta$Post.Prob)`. The heterogeneity of effect sizes was $\sigma_{\text{experiment}}^2 = `r summ_meta$sigma2_experiment`$, 95% CrI [`r summ_meta$sigma2_experiment.lower`, `r summ_meta$sigma2_experiment.upper`], at the experiment level and $\sigma_{\text{article}}^2 = `r summ_meta$sigma2_article`$, 95% CrI [`r summ_meta$sigma2_article.lower`, `r summ_meta$sigma2_article.upper`] at the article level. Therefore, `r print_perc(1 - summ_meta$icc)` of the heterogeneity between effect sizes was attributable to differences between experiments *within* articles and `r print_perc(summ_meta$icc)` was attributable to differences *between* articles.

@@

As a next step, we conducted a meta-regression analysis to test if the gender of the infants, their age, or the type of mental rotation task was related to mental rotation performance (Figure\ \@ref(fig:fig2)). Indeed, experiments with all-male samples revealed larger looking time differences than experiments with all-female samples, $b = `r summ_reg$male_female`$ (where $b = \Delta{g}$), 95% CrI [`r summ_reg$male_female.lower`, `r summ_reg$male_female.upper`]. The probability for this effect being larger than zero was `r print_prob(hyps_brm_reg$male_female$Post.Prob)`. We found no difference between mixed-gender and all-female samples, $b = `r -summ_reg$female_mixed`$, 95% CrI [`r -summ_reg$female_mixed.lower`, `r -summ_reg$female_mixed.upper`] (`r print_prob(hyps_brm_reg$female_mixed$Post.Prob)` probability of an effect larger than zero). Additionally, mean age was not related to mental rotation performance, with a change per year of $b = `r summ_reg$age`$, 95% [`r summ_reg$age.lower`, `r summ_reg$age.upper`]. We also did not detect an interaction between gender and age ([females - mixed] $\times$ age: $b = `r summ_reg$female_mixed_age`$, 95% CrI [`r summ_reg$female_mixed_age.lower`, `r summ_reg$female_mixed_age.upper`]; [males - females] $\times$ age: $b = `r summ_reg$male_female_age`$, 95% CrI [`r summ_reg$male_female_age.lower`, `r summ_reg$male_female_age.upper`]). Finally, there was weak evidence that violation of expectation tasks yielded larger effects than habituation tasks, $b = `r summ_reg$voe_habituation`$, 95% [`r summ_reg$voe_habituation.lower`, `r summ_reg$voe_habituation.upper`]. The probability for this effect being greater than zero was `r print_prob(hyps_brm_reg$voe_habituation$Post.Prob)`.

@@

# Combine the two regression plots
plot_grid(
  plot_reg, plot_coef,
  nrow = 1, rel_widths = c(3, 2), labels = "AUTO",
  label_size = 11, label_fontfamily = "Helvetica",
  label_x = 0.008, label_y = 0.99
) +
  draw_plot(plot_reg_legend, x = 0.385, y = 0.865, hjust = 0.5, vjust = 0.5)

@@

(ref:fig2-caption) **Effects of gender, age, and task.** (A) Squares show the effect size (Hedges' $g$) for infants' mental rotation performance in each of the 59 individual experiments. Squares are color-coded according to the type of habituation task and the gender of the infants (blue = habituation task, all-female sample, yellow = habituation task, all-male sample, red = habituation task, mixed-gender sample, purple = violation of expectation [VoE] task, mixed-gender sample). Lines indicate the best-fit regression estimates according to a Bayesian three-level meta-regression model and gray ribbons indicate their corresponding 50% and 95% credible interval (CrI). (B) Fixed effect estimates obtained from the Bayesian three-level meta-regression model are depicted as black dots together with their 50% CrI (thick black lines) and 95% CrI (thin black lines). Gray curves indicate the posterior distribution for each effect.

@@

To confirm the gender difference between males and females, we set up another Bayesian meta-analysis, this time focusing on the looking-time contrasts between male and female infants reported *within* each experiment by the original authors. Our additional analysis revealed a meta-analytic effect size of $g = `r summ_gender$intercept`$, 95% CrI [`r summ_gender$intercept.lower`, `r summ_gender$intercept.upper`] (Supplementary Figure\ \@ref(fig:figs1)) and a probability of this effect being greater than zero of `r print_prob(hyp_gender$Post.Prob)`. The heterogeneity of effect sizes was $\sigma_{\text{experiment}}^2 = `r summ_gender$sigma2_experiment`$, 95% CrI [`r summ_gender$sigma2_experiment.lower`, `r summ_gender$sigma2_experiment.upper`], at the experiment level and $\sigma_{\text{article}}^2 = `r summ_gender$sigma2_article`$, 95% CrI [`r summ_gender$sigma2_article.lower`, `r summ_gender$sigma2_article.upper`] at the article level. Therefore, `r print_perc(1 - summ_gender$icc)` of the heterogeneity between effect sizes was attributable to differences between experiments *within* articles and `r print_perc(summ_gender$icc)` was attributable to differences *between* articles.

@@

We inspected funnel plots and performed Egger regression tests to examine the possibility of publication bias in the literature included here. For the meta-analysis of mental rotation performance (Figure\ \@ref(fig:fig3)a), we observed a slight asymmetry in the funnel plot, indicating a small publication bias. This was confirmed by an Egger regression test that showed a reliable association between effect sizes and their corresponding standard errors, $b = `r coef(egger_meta$fit)["Xsei"]`$, $t(`r egger_meta$dfs`) = `r egger_meta$zval`$, $p = `r print_p(egger_meta$pval)`$, 95% confidence interval (CI) [`r egger_ci_meta["2.5 %"]`, `r egger_ci_meta["97.5 %"]`] (two-sided test). Nevertheless, a jackknife (leave-one-out) analysis confirmed that the current results are robust to the effects of individual outlier experiments (Supplementary Table\ \@ref(tab:tabs3)). For the meta-analysis of gender differences in mental rotation performance (Figure\ \@ref(fig:fig3)b), the asymmetry in the funnel plot was less pronounced and the slope of the Egger regression test was not statistically significant, $b = `r coef(egger_gender$fit)["Xsei"]`$, $t(`r egger_gender$dfs`) = `r egger_gender$zval`$, $p = `r print_p(egger_gender$pval)`$, 95% CI [`r egger_ci_gender["2.5 %"]`, `r egger_ci_gender["97.5 %"]`] (two-sided test).

@@

# Combine plots and legend
plot_grid(
  plotlist = plots_funnel, nrow = 1, labels = "AUTO",
  label_size = 11, label_fontfamily = "Helvetica"
) +
  draw_plot(legend_funnel, x = -0.39, y = 0.355)

@@

(ref:fig3-caption) **Evaluation of publication bias.**  Funnel plots for the meta-analysis of mental rotation performance (A) and for the meta-analysis of gender differences in mental rotation performance (B). The plots show the standard error and effect size for each of the individual experiments as a black square. The funnel contours (diagonal black lines) depict a 95% pseudo-confidence interval around the meta-analytic effect size (vertical black line). Gray shades indicate a 95% pseudo-confidence interval (dark gray) and a 99% pseudo-confidence interval (light gray) under the null hypothesis. These shades thus illustrate which of the original experiments observed a significant effect. For the meta-analysis of mental rotation performance, a slight asymmetry induced by the underrepresentation of experiments with high standard errors and small effect sizes suggests a small publication bias.

@@

To facilitate the interpretability of our models, the effect size of the looking time difference between novel and familiar rotation events was coded in a linear fashion. Accordingly, a positive looking time difference (i.e., a novelty preference) was taken as stronger evidence for mental rotation than a looking time difference of zero (i.e., no systematic preference). A zero difference in turn was taken as stronger evidence for mental rotation than a negative looking time difference (i.e., a familiarity preference). This approach can be corroborated by the established view that novelty preference is generally considered as the paradigmatic mental rotation behavior [for related discussion, see @black2017; @cristia2018; @houston-price2004]. Nevertheless, some authors have argued that familiarity preference should also be taken as evidence for mental rotation ability [@erdmann2018; @moore2020; @johnson2020]. Interpreting novelty preference *and* familiarity preference as successful mental rotation would have increased the effect sizes reported here.

@@

A number of additional factors have been associated with individual differences in infant mental rotation performance. For example, mental rotation is related to previous relevant experience with the particular objects used in the specific task [@mohring2013; @schwarzer2013a; @slone2018]. This relation also applies to previous experience with manually rotating toys [@schwarzer2013a]. While these preliminary results require replication, they are in line with the longstanding notion that prior knowledge is the strongest predictor of learning outcomes in a range of cognitive domains [e.g., @ausubel1968; @bradley1983; @halberda2008]. Furthermore, there is yet to be confirmed preliminary evidence for possible links between mental rotation performance and several sensory motor skills including fine and gross motor skills, oculomotor control, and crawling skills [@schwarzer2013; @schwarzer2013a].

@@

# Method

@@

In the present meta-analysis, we followed the established PRISMA 2020 guidelines [Preferred Reporting Items for Systematic reviews and Meta-Analyses\; @page2021]. The PRISMA checklist is provided in Supplementary Table\ \@ref(tab:tabs4) while Supplementary Figure\ \@ref(fig:figs2) displays the PRISMA flowchart. We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

@@

We entered the search terms ("mental rotation" OR "mental transformation" OR "spatial rotation" OR "spatial transformation" OR "spatial ability" OR "spatial skills") AND ("infant" OR "infants" OR "infanthood" OR "toddler" OR "toddlers" OR "toddlerhood" OR "child" OR "children" OR "childhood" OR "month" OR "months") into four online databases (APA PsycINFO, PubMed/MEDLINE, Scopus, and ProQuest Dissertations & Theses Global). All database queries were completed on December 6, 2021. We configured the databases to check for article titles, abstracts, and keywords while applying no other filters or limits. This yielded 2,616 articles in total, 1,954 of which remained after removing duplicate records (Supplementary Figure\ \@ref(fig:figs2)). We further identified 76 articles by screening the reference sections of previous reviews and meta-analyses on mental rotation and related skills [@frick2014; @johnson2020; @kubicek2018; @lauer2019a; @linn1985; @moore2020; @uttal2013; @voyer1995; @yang2020]. Of these, 34 articles had not been covered by the database search. We also identified 94 articles by screening the reference sections of all publications that had been included after the first pass of the selection process. Of these, 49 articles had not been covered by the database search. Accordingly, we screened 2,037 unique articles in total.

@@

The final sample size of articles was determined by including all articles that fulfilled the inclusion criteria described above. Two independent raters read the abstract and, if necessary, relevant sections of the full text to check if an article fulfilled the inclusion criteria. Interrater agreement for the binary decision to include versus exclude an article was `r print_perc(irr_bin_perc)` ($\kappa_w$ [Cohen's weighted kappa] $= `r irr_bin_kappa_w$estimate`$, 95% CI [`r irr_bin_kappa_w$lower`, `r irr_bin_kappa_w$upper`]). Interrater agreement for the specific eligibility criteria was `r print_perc(irr_code_perc)` ($\kappa_w$ $= `r irr_code_kappa_w$estimate`$, 95% CI [`r irr_code_kappa_w$lower`, `r print_num(irr_code_kappa_w$upper)`]). Cases where the two ratings diverged were resolved via discussion among all raters until a consensus was reached. One article [@mash2007] was excluded because the authors used a unique mental rotation paradigm that was not comparable to the paradigms used in the other articles. Another article [@pedrett2020] was excluded because the average age of the infants studied (`r print_num(descs_pedrett2020$age_months, digits = 1)` months) was almost twice as high as the average age of the next article (`r print_num(descs_pedrett2020$next_largest_age_months, digits = 1)` months; $z = `r descs_pedrett2020$z`$ compared to all articles). We therefore decided to narrow our analysis from the first 3 years of life to the first 16 months of life. This procedure led to a total of `r descriptives$n_articles` articles being included in the meta-analysis (Supplementary Figure\ \@ref(fig:figs2)).

@@

We verified the results obtained from our Bayesian analyses using classical frequentist meta-analysis and meta-regression. To this end, we used the metafor package [Version `r packageVersion("metafor")`\; @viechtbauer2010] in R to specify the same three-level models as described above but without the Bayesian priors (Supplementary Table\ \@ref(tab:tabs2)). These models were fitted using restricted maximum likelihood estimation (REML). To verify that this procedure converged on the most probable estimates, we examined profile likelihood plots [@raue2009; @viechtbauer2010] for the two variance components in the model (i.e., the random effects of experiments and articles; Supplementary Figure\ \@ref(fig:figs4)).
